Knowing your role: Varying expectations in software organizations

### 2011-09-27 soulgem.tex
Third draft, Study Design and Execution.

\section{Method and Study Execution}

\subsection{Study Design}

We performed a qualitative empirical study based on the theoretical framework presented above to improve our understanding of roles in software organizations. We chose to follow Yin's case study methodology \cite{Yin2003}. We executed an explanatory two-case case study with two embedded units of analysis (software organizations and roles).

We formulated eight research questions to guide our data collection and analysis. For each of them, we also formulated a corresponding proposition based on the theoretical framework we delineate above.

@@

\emph{Proposition:} We do not expect to be able to provide a simple classification of roles. However, we hypothesize that some expectations often come together, and that these patterns will form a partial picture of roles in software development.

@@

\textbf{\emph{RQ3:}} What is the logic (what are the factors) behind the particular clusters of expectations found in software organizations?

@@

\emph{Proposition:} There is a significant divergence between roles and positions, and the reason is that position is an institutional factor, as specified above, and there are many other factors that contribute to the shaping of a role.

\textbf{\emph{RQ5:}} How do roles change over time in a software organization?

@@

\subsection{Study Execution and Analysis}

We collected field data from two software organizations in the area of Victoria, Canada. Our case selection was opportunistic: we wanted groups that develop software as their main activity, and whose headquarters were located in our vicinity. We contacted managers in both organizations and negotiated access to their sites in return for feedback on their development practices.

We carried out our case study during the Winter and Spring of 2011. The first author spent five weeks performing observations, three weeks at the first and two at the second site, sitting at a desk in their development areas, attending to several kinds of meetings (stand-up, product, all-hands, and others), and exploring their issue tracking and project estimation systems.

@@

Following Yin's \cite{Yin2003} suggestions, we analyzed all the observation and interview data examining whether it supported or challenged our study propositions. We were particularly interested in evidence that did not fit with our theoretical framework, in order to refine it for future studies.

>>>>

\section{Study Design and Execution}

We performed an empirical study based on our theoretical framework to improve our understanding of roles in software organizations. We chose to follow Yin's case study methodology \cite{Yin2003}. We executed an explanatory two-case case study with two embedded units of analysis (software organizations and roles).

We formulated six research questions to guide our data collection and analysis. For each of them, we formulated a corresponding proposition of what we expected to find.

@@

\emph{Proposition:} We do not expect to be able to provide a simple classification of roles. However, some expectations often come together, and these patterns may form a partial picture of roles in software development.

@@

\textbf{\emph{RQ3:}} What is the logic behind the particular clusters of expectations found in software organizations?

@@

\emph{Proposition:} There is a significant divergence between roles and positions, and the reason is that \emph{position} describes at best a subset of the expectations shaping a role.

\textbf{\emph{RQ5:}} How do roles change over time in software groups?

@@

We collected field data from two software organizations in the area of Victoria, Canada. Our case selection was opportunistic: we wanted software development groups whose headquarters were in our vicinity. We negotiated access to both sites in return for feedback on their development practices.

The first author spent five weeks performing field observations, three weeks at the first site and two at the second, sitting at a desk in their development areas, attending to several kinds of meetings, and exploring their issue tracking and project estimation systems.

@@
